Commutative Character Sheaves, with an application to generic types for supercuspidal representations

### 2015-09-06 followup.tex
Diagrams added to descent to abelianization
Commit before talking to David

Let $G$ be a smooth group scheme.
Let $j: G\der \hookrightarrow G$ denote the closed subgroup generated by the commutator morphism $c: G\times G \to G$ given by $c(x,y) = xyx^{-1}y^{-1}$; see \cite{SGA3}*{{VI}$_B$, 7.2}.
Let $n : G\der \times G \to G$ be the restriction of $m: G\times G \to G$.

@@

@@

\begin{lemma}\label{lem:smooth_coeq}
The quotient $q : G \to G\ab\ceq G/G\der$ is a smooth $G\der$-torsor and is the coequalizer of $n : G\der \times G \to G$ with projection $p: G\der \times G \to G$. 
\[
\begin{tikzcd}
G\der\times G
 \arrow{r}[shift left]{n}
  \arrow{r}[shift right,swap]{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]
\end{lemma}

@@

\section{Geometrization of generic types for supercuspidal representations}

Let $K$ be a local field with residue field $\Fq$.
Let $\underline{G}$ be a smooth group scheme over $K$.
Set $G_n = \Gr^R_n(\underline{G})$; this is a smooth group scheme over $\Fq$.
Set $G = \Gr^R_n(\underline{G})$; this is a group scheme over $\Fq$ equipped with $G = \varprojlim_{n\in \NN} G_n$.
Let $\QCS(\underline{G})$ be the category of commutative character sheaves on $G$. . . .  as in \cite{cunningham-roe:13a}.

\begin{corollary}\label{cor:QCS}
$\Hom(\underline{G}(\OK),\EEx) \iso \QCSiso{\underline{G}}$.
\end{corollary}

>>>>

Let $G$ be a smooth group scheme.
Let $j: G\der \hookrightarrow G$ denote the closed subgroup generated by the commutator morphism $c: G\times G \to G$ given by $c(x,y) = xyx^{-1}y^{-1}$; see \cite{SGA3}*{{VI}$_B$, 7.2}.

\[
\begin{tikzcd}
{} & G\times G \arrow{d}{c} &&&\\
1 \arrow{r} & G\der \arrow{r}{j} & G \arrow{r}{q} & G\ab \arrow{r} & 1
\end{tikzcd}
\]

@@

Let $n : G\der \times G \to G$ be the restriction of $m: G\times G \to G$.

@@

The quotient $q : G \to G\ab$ with kernel $G\der$  is the coequalizer of $n : G\der \times G \to G$ with projection $p: G\der \times G \to G$. 
\[
\begin{tikzcd}
G\der\times G
 \arrow{r}[shift left]{n}
  \arrow{r}[shift right,swap]{p}
&
G 
\arrow{r}{q}
& 
G\ab
\end{tikzcd}
\]

\begin{lemma}\label{lem:smooth_coeq}
The quotient $q : G \to G\ab\ceq G/G\der$ is a smooth $G\der$-torsor.
\end{lemma}

[I don't have a proof of this yet.]

\[
\begin{tikzcd}
\CCS(G\der \times G)
& \arrow{l}[shift left,swap]{n^*}
  \arrow{l}[shift right]{p^*}
\CCS(G) 
&&\arrow{ll}[swap]{f^*} 
\CCS(G\ab) \arrow{dl}{\text{`comparison'}}\\
& & \arrow{ul}{\text{forget}} \CCS_{G\der}(G) & 
\end{tikzcd}
\]

@@

\section{Application to certain supercuspidal representations of $p$-adic groups}

[A basic question which I think we must answer but which we can maybe skate around for now:
When is the Suzuki transform of a K-variety representable as a k-scheme?

Let $K$ be a local field with residue field $\Fq$.
Let $\underline{G}$ be a smooth group scheme over $K$.
Set $G_n = \Gr^R_n(\underline{G})$; this is a smooth group scheme over $\Fq$.
Set $G = \Gr_R(\underline{G})$; this is a group scheme over $\Fq$ equipped with $G = \varprojlim_{n\in \NN} G_n$.
Let $\QCS(G)$ be the category of commutative character sheaves on $G$. . . .  as in \cite{cunningham-roe:13a}.

\begin{proposition}\label{prop:QCS}
$\Hom(\underline{G}(\OK),\EEx) \iso \QCSiso{G}$.
\end{proposition}
